Mechanisms of detachment in fibrillar adhesive systems

### 2021-06-09 source/tex/manuscript.tex
finalized MS for submission to acta
fixed references format
fixed hinge/joint phrase
included statement of significance

\author[mymainaddress]{Michael Kappl\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{kappl@mpip-mainz.mpg.de}

@@

@@

In the energy balance approach, adhesion is usually characterized by ``work
of adhesion'' (W\textsubscript{adh}), which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, which increases W\textsubscript{adh}
and thus adhesion is enhanced \cite{RN29,RN114}. Detachment of an
individual hair can be explained based on Kendall's peeling theory \cite{RN147, RN296},
which predicts low forces at high peeling angles.

In the force balance approach, adhesion is characterized by pull-off
force, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a ``cohesive zone model'', Hui et. al. \cite{RN72}
had identified two regimes of single hair detachment: 1) a \emph{flaw sensitive}
regime, where, for large hair raidus, contact failure occurs due to crack propagation, 
initiated by a stress singularity at the edge of the hair, leading
to low \textgreek{sv}\textsubscript{p}; 2) a \emph{flaw insensitive} regime,
where, for small hair raidus, the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Tian et. al. \cite{RN230} had shown
that the ``spatula'' shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. Federle \cite{RN20}
had further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally, and easily detached by elastic
recoil when pushed distally.

@@

The fibrillar adhesive pad is assumed to be a one dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg, assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an instantaneous angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net force on the pad and $M_{net}$ is the the net
torque about the hinge, at a particular instant during the detachment
process.

\begin{figure}[H]
\includegraphics{Figure1-model_schematic}\caption{Spring contact model of a fibrillar adhesive pad. The pad consists of
an array of $N_{t}$ hairs connected to a deformable leg at the hinge.
At a particular distance, $d_{s}$, $n$ number of hairs are in contact
and the array is oriented at a tilt angle, $\theta$, with the surface.
\label{fig:Schematic}}
\end{figure}

@@

Simplifying, we get (see Appendix for derivation):

@@

The net torque, $M_{net}$, about the hinge can be similarly derived
(see Appendix):

@@

Let us now consider the scenario where even the leg above the hinge
can undergo elastic stretching together with the hairs. When a hair
detaches from the surface, the leg undergoes an elastic recoil due
to the stored elastic energy. Suppose the leg relaxes upward by a
recoil length, $\Delta l$. For $n$ hairs in contact, the force balance
before and after a hair detaches is given respectively by:

@@

We consider three possible mechanisms to detach the adhesive pad from
a surface: 1) \emph{Fixed pull}, where the pad is pulled vertically
up while keeping a fixed leg-hinge, 2) \emph{Free pull,} where the
pad is pulled vertically up while keeping the leg-hinge free to allow
rotation of the array and 3) \emph{Flex}, where the pad is hinged
to an external point (claw-hinge), and detached in a rotary fashion,
emulating the claw function in insects. To investigate each case in
detail, let us assume a pad with $N_{t}=25$ hairs and $\hat{f_{p}}=0.1$
attached to a stiff leg ($k_{l}\rightarrow\infty$). The situation
of a soft leg ($k_{l}/k_{h}=10$) is also considered for the first
two cases involving vertical detachment.

@@

\begin{figure}[H]
\includegraphics{Case1-Fixed_pull}\caption{Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a fixed hinge between the pad and the leg (\emph{Fixed pull}). The
tilt angle, $\theta$, of the array is kept fixed during detachment.
The leg is either stiff ($k_{l}\rightarrow\infty$) or soft ($k_{l}/k_{h}=10$).
Positive force values represent attraction. The green dashed line
represents the maximum possible adhesion for the pad. All values are
normalized to dimensionless forms, as described in text.\label{fig:Fixed-pull}}
\end{figure}

The fibrillar adhesive pad can be detached by pulling it vertically upwards
while maintaining a constant tilt angle, $\theta$, with the surface.
This can be achieved if the joint connecting the leg to the array
(leg-hinge) is kept fixed. Equations \ref{eq:force_n_d}, \ref{eq:d_max}
and \ref{eq:simul_detach} can be used to get the resulting force-distance
curves.

@@

\includegraphics{Case2-Free_pull}\caption{Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a free hinge between the pad and the leg (\emph{Free pull}). $s$
is the distance between the leg-hinge and the right end of the array
and $L$ is the array length. The free hinge allows further tilting
of the array during the vertical pull. The leg is either stiff ($k_{l}\rightarrow\infty$)
or soft ($k_{l}/k_{h}=10$). Positive force values represent attraction.
The green dashed line represents the maximum possible adhesion for
the pad. All values are normalized to dimensionless forms, as described
in text. \label{fig:Free-pull}}

@@

Similar to the previous case, we once again consider the situation
where the adhesive pad is pulled vertically upwards for detachment.
However now, the leg-hinge is assumed to be free. In this case, the
array will reorient itself to maintain a zero net torque about the
leg-hinge during the entire detachment process. At any given instant,
the tilt angle, $\theta$, can be found by setting $M_{net}$ to zero
in equation \ref{eq:moment_n_d} to get:

@@

Using the above relation together with equations \ref{eq:force_n_d},
\ref{eq:d_max} and \ref{eq:simul_detach}, we can find force-distance
curves during a free vertical pull of the adhesive pad. Since the
position of the leg-hinge will influence the net torque, we use the
ratio, $s/L$, to study its effect on the detachment forces.

Maximum adhesion is seen when the leg-hinge is positioned at the centre
of the array, i.e. $s/L=0.5$ (Figure \ref{fig:Free-pull}). Here,
the net torque due to the hairs is balanced by symmetry and the array
remains parallel to the substrate until all hairs detach simultaneously
at $\hat{d_{s}}=0.1$. Shifting the position of the leg-hinge further
away from the array centre leads to lower forces or adhesion. The
resulting torque imbalance will tilt the array, which reduces the
net force due to the \emph{elastic weakening} effect, as described
in the previous section. Higher $s/L$ increases the net torque to
be balanced, leading to a higher tilt of the array and thus lower
net force.

@@

\begin{figure}[H]
\includegraphics{Case3-Fixed_flex}\caption{Force curves for a fibrillar adhesive pad detached by flexing it about the claw.
$\hat{F}_{pull}=\frac{F_{pull}}{k_{h}w}$ is the normalised pulling
force necessary to apply the torque about the claw-hinge for detachment,
$\hat{F}_{hinge}=\frac{F_{hinge}}{k_{h}w}$ is the normalised reaction
force on the claw-hinge, $\hat{d_{s}}=\frac{d_{s}-l_{h,0}}{w}$ is
the normalized vertical distance of the claw-hinge from the surface.
Here, $s_{l}/L=1$ and $s_{h}/w=10$. The green dashed line represents
the maximum possible adhesion for the pad. \label{fig:Flex}}
\end{figure}

Instead of a vertical pull, the adhesive pad can also be detached
by rotating it about the claw-hinge, located outside the array. Such
a mode of detachment will be driven by a torque applied by the leg
to rotate the pad around the claw-hinge until all the hairs lose contact.
Let $s_{h}$ be the distance between the claw-hinge and the right
end of the array; $s_{l}$ be the distance between the fixed leg-hinge
and the right end of the array. To illustrate the mechanism, let us
fix $s_{l}/L=1$ and $s_{h}/w=10$ and vary the vertical claw-hinge
distance, $d_{s}$. At any particular instant, the pulling force applied
by the leg, $F_{pull}=M_{net}/\left(s_{l}+s_{h}\right)$, where $M_{net}$
is obtained by setting $s=-s_{h}$ in equation \ref{eq:moment_n_d}.
Equation \ref{eq:force_n_d} will give us the reaction force acting
on the claw-hinge, $F_{claw-hinge}$.

@@

The dimensionless parameter, $\hat{f}_{p}=\frac{f_{p}}{k_{h}w}$,
governs the strength and compliance of the array, where, high values
represent a dense array of strongly adhering soft hairs. Let us consider
the case of an adhesive pad with $N_{t}=25$ hairs and look at how
$\hat{f}_{p}$ influences the reduction factor under each mode of
detachment (Figure \ref{fig:Reduction-factor_fp}). From the heat
maps, one can generally say that lower values of $\hat{f}_{p}$ increase
the reduction factor and thus provide a greater possibility of controlling
adhesion.

\begin{figure}[H]
\noindent \includegraphics{Reduction_factor-fp}\caption{Heat maps showing the effect of the dimensionless parameter, $\hat{f}_{p}$,
on the reduction factor for each mode of detachment. Here, we fix
the number of hairs, $N_{t}=25$. All values are normalized to dimensionless
forms, as described in text. \label{fig:Reduction-factor_fp}}
\end{figure}

@@

For the case of detachment via \emph{free pull}, $\hat{f}_{p}$ has
no influence on the reduction factor. On the other hand, shifting
the position of the leg-hinge further away from the array (i.e. high
$s/L$) results in large values of $r$. In this scenario, the higher
torque exerted by the array leads to a higher tilt, and thus increases
the reduction factor via \emph{elastic weakening}, saturating to the
maximum value of 25.

@@

Let us now fix $\hat{f}_{p}=0.1$ and investigate the influence of
the number of hairs, $N_{t}$, on the reduction factor (Figure \ref{fig:Reduction-factor-Nt}).
The heat maps show that high $N_{t}$ increases $r$ irrespective
of the mode of detachment. Under a tilted state, more number of hairs
are compressed when $N_{t}$ is high, which reduces the net adhesion.
This highlights the advantage of having a split contact design found
in many biological systems. Higher number of hairs offer a better
control over adhesion and thus is more suited for reversible attachment
and detachment during locomotion. The specific trends of reduction
factor for each mode of detachment can be understood by similar arguments
of \emph{elastic weakening}, as discussed in the previous section.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics{Reduction_factor-Nt}\caption{Heat maps showing the effect of the number of hairs, $N_{t}$, on
the reduction factor for each mode of detachment. Here, we fix the
dimensionless parameter, $\hat{f}_{p}=0.1$. All values are normalized
to dimensionless forms, as described in text. \label{fig:Reduction-factor-Nt}}
\par\end{centering}
\end{figure}

@@

Based on the above assumptions, we can come up with force-distance
curves to detach the beetle's leg via various pathways (Figure \ref{fig:Detachment-pathways}).
First, let us assume the hinge H\textsubscript{2} to be fixed, such
that both the distal and proximal pads are combined to be a single
long pad with $N_{t}=60$ hairs. Path 1 shows the case where the pad
shows maximum possible adhesion. Here, the combined pad is vertically
pulled upwards while keeping the array perfectly parallel to the surface.
If the pad is detached by keeping H\textsubscript{1} fixed and maintaining
a tilt of $1\lyxmathsym{\textdegree}$ with the surface (path 2),
the forces dramatically reduces, with around 10 times reduction in
the adhesion compared to path 1. We can also detach the pad by switching
between the different mechanisms. Path 3 shows one such example, where,
initially the leg is pulled vertically up while keeping H\textsubscript{1}
fixed, stretching the hairs. On reaching point $a$, H\textsubscript{1}
is set free, which results in a sudden drop in force due to the excess
torque by the stretched hairs, tilting the array. Beyond this, the
force curve follows the free pull mechanism, with $\thicksim3.5$
times reduction in adhesion. An alternate strategy of switching between
mechanisms would be to first apply a load on the pad (path 4) and
compress the hairs until point $b$. Beyond this point, the claw can
be used as a hinge to detach the pad via flexing it around H\textsubscript{3},
which once again reduces the adhesion force. Now, if we assume the
hinge H\textsubscript{2} to be free such that the two pads can behave
distinctly, we can consider the scenario where the proximal pad is
flexed around the distal pad at H\textsubscript{2} while keeping
H\textsubscript{1} fixed (path 5). After the proximal pad has completely
detached, H\textsubscript{1} can be freed up at point $c$ to detach
the distal pad via \emph{free pull} with very little force. This pathway
results in a $\thicksim5$ times reduction in adhesion.

\begin{figure}[H]
\includegraphics{Detachment_pathways}\caption{Force distance curves showing possible detachment pathways for a dock
beetle's leg, derived from the spring contact model. The curves are
offset laterally for clarity. Colours represent the distinct detachment
pathways, as labelled correspondingly, with arrows indicating the
direction of retraction. The different line style denotes the specific
detachment mechanism followed by any region of the pathway. The inset
schematic shows the assumed locations of the different hinges (H\protect\textsubscript{1},
H\protect\textsubscript{2} and H\protect\textsubscript{3}) employed
by the leg.\label{fig:Detachment-pathways}}
\end{figure}

The above analysis illustrates how the design of the beetle's hairy
adhesive pads is suitable for modulating its adhesion. Effective control
and release of its leg joints can help the insect to reduce the pad's
adhesion, allowing it to detach with little effort. High reduction
in adhesion is seen when the pad is tilted relative to the surface
during detachment, as a result of \emph{elastic weakening}. To the
best of our knowledge, there is no direct experimental evidence that
beetles can modulate its adhesion by taking advantage of this effect.
Considering that hair deformation occurs at length scales below 10
\textgreek{m}m, direct observation of this effect on running beetles
would be challenging. A recent study on PDMS micro-pillar arrays,
however, does indeed show a strong reduction in the adhesion force
due to slight misalignments with the surface \cite{RN252}. Based on
previously reported microscopic investigation of freely walking dock
beetles \cite{RN83}, we argue that the following experimental observations
provide support to our proposed theory: 1) The detachment was shown
to follow a three-dimensional twist of the leg, which suggests a complex
inhomogeneous deformation of hairs across the array, leading to \emph{elastic
weakening}. 2) The beetle can at times instantaneously detach all
its legs and ``drop'' itself while upside down. This could be explained
by the beetle freeing up its leg joints and using just its body weight
to provide the necessary force to detach all its legs via \emph{free
pull} (similar to path 3 above). 3) Only a fraction of the beetle's
pads made contact with the surface during locomotion, which indicates
that the pads should naturally be in a slightly tilted state. This
not only reduces the contact area, but also non-uniformly deforms
the hairs, both leading to a reduction in adhesion for easy detachment.
4) Contact images showed that the array ``peels'' from the proximal
to distal direction during detachment. However, the beetle's hairs
are attached to a relatively stiff backing \cite{RN53}, so it wouldn't
be able to ``peel'' its array, since peeling, strictly speaking,
depends on the elastic contribution of a thin flexible backing as
it bends during the process \cite{RN147}. Rather, the ``peeling''
observed in the beetle's case should be a result of the pad detaching
from the surface in a tilted orientation, causing the hairs to detach
in sequence. 5) The time scale of detachment was reported to be an
order of magnitude shorter than the attachment time scale, which could
be a result of the elastic recoil of the leg causing a spontaneous
propagation of the detachment front (Figure \ref{fig:Fixed-pull}).

There exists a limit to how much the pad can tilt, depending on its
geometry and material properties. Suppose the hair has a maximum linear
elastic strain limit, $\varepsilon_{m}$, and natural length, $l_{h,0}$.
Based on Figure \ref{fig:Schematic}, if the right most hair is compressed
to its elastic limit, we can show from simple geometry, that, the
corresponding maximum limit in tilt angle is given by: 
\[
\theta_{limit}=\arctan\frac{l_{h,0}\varepsilon_{m}}{\left(N_{t}-1\right)w}
\]
$\theta_{limit}$ will limit the reduction factor for each of the
detachment mechanisms presented. Longer hairs can result in a lateral
collapse or bundling of hairs, imposing an additional constraint on
$\theta_{limit}$. Large deformation of hairs can also lead to buckling,
which will further limit the reduction in adhesion due to the smaller
effective modulus. Thus, the geometry of the individual hairs is also
crucial in effectively reducing adhesion by \emph{elastic weakening}.
Ideally, hairs with large diameters would be preferable to eliminate
buckling, and optimally long hairs would provide a large range of
angles to which the pad can be tilted while avoiding bundling.

@@

The net torque about the hinge due to the deformed hairs of the array
is,

@@

Here, $\lambda_{i}=s-\left(\frac{n-1}{2}w-\frac{x_{i}}{\cos\theta}\right)$
is the length of the lever arm between the $i^{th}$ hair and the
hinge.

>>>>

\author[mymainaddress]{Michael Kappl\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author. \emph{Telephone:} +49 6131 379-114}
\ead{kappl@mpip-mainz.mpg.de}

@@

\section*{Statement of Significance}
Insects such as flies or beetles have the ability to walk on any surface against gravity. It was previously shown that
these insects possess a dense array of micrometer sized hairs on their legs. These hairy legs are responsible for the 
insect's strong adhesion to surfaces by resisting contact failure due to micro-roughness or defects on the surface.
In this work, we theoretically show that, a hairy design of the leg not only provides strong adhesion, but can also allow easy detachment from a surface. This is achieved simply by modulating the leg's tilt relative to the surface with the aid of its leg joint or claws. Such a design can thus inspire artificial adhesives with switchable adhesion properties.

@@

In the energy balance approach, adhesion is usually characterized by ``work
of adhesion'' (W\textsubscript{adh}), which is the energy required
to separate a pad from the surface. During detachment, the elastic
energy stored in the hair is dissipated, which increases W\textsubscript{adh}
and thus adhesion is enhanced \cite{RN29,RN114}. Detachment of an
individual hair can be explained based on Kendall's peeling theory \cite{RN147, RN296},
which predicts low adhesion at high peeling angles.

In the force balance approach, adhesion is characterized by pull-off
force, F\textsubscript{p} (or stress, \textgreek{sv}\textsubscript{p}),
which is the minimum force necessary to separate two surfaces from
contact. Based on a ``cohesive zone model'', Hui et. al. \cite{RN72}
had identified two regimes of single hair detachment: 1) a \emph{flaw sensitive}
regime, where, for large hair radius, contact failure occurs due to crack propagation, 
initiated by a stress singularity at the edge of the hair, leading
to low \textgreek{sv}\textsubscript{p}; 2) a \emph{flaw insensitive} regime,
where, for small hair radius, the contact interface fails simultaneously, leading to high
\textgreek{sv}\textsubscript{p}. Tian et. al. \cite{RN230} had shown
that the ``spatula'' shaped hair tips in a gecko's toe allows it
to change adhesion by three orders of magnitude by laterally sliding
and controlling the pulling angle to disorient the hairs. Federle \cite{RN20}
had further argued that the curved shape of the hair helps the pad
to stay attached when pulled proximally, and easily detached by elastic
recoil when pushed distally.

@@

The fibrillar adhesive pad is assumed to be a one dimensional array of $N_{t}$
hairs, each behaving like a spring with spring constant, $k_{h}$,
and natural length, $l_{h,0}$ (Figure \ref{fig:Schematic}). The
array backing is assumed to be stiff. The pad is attached to a linearly
deformable leg, assumed to be another spring with spring constant,
$k_{l}$, and natural length, $l_{l,0}$. The leg is hinged to the
array at a distance, $s$, from the right end of the array. The hinge (or joint) 
is at a vertical distance, $d_{s}$, from the surface. The hairs are
spaced apart by a width, $w$, and the array is of length, $L=\left(N_{t}-1\right)w$.
The pad is oriented at an instantaneous angle, $\theta$, while making contact with
a flat smooth surface. Each hair can attain a maximum length, $l_{h,p}$,
before pull-off, such that its pull-off force, $f_{p}=k_{h}\left(l_{h,p}-l_{h,0}\right)$.
$F_{net}$ is the net force on the pad and $M_{net}$ is the the net
torque about the joint, at a particular instant during the detachment
process.

\begin{figure}[H]
\includegraphics{Figure1-model_schematic}\caption{\textbf{Spring contact model of a fibrillar adhesive pad.} The pad consists of
an array of $N_{t}$ hairs connected to a deformable leg at the joint.
At a particular distance, $d_{s}$, $n$ number of hairs are in contact
and the array is oriented at a tilt angle, $\theta$, with the surface.
\label{fig:Schematic}}
\end{figure}

@@

Simplifying, we get (see Appendix A for derivation):

@@

The net torque, $M_{net}$, about the joint can be similarly derived
(see Appendix A):

@@

Let us now consider the scenario where even the leg above the joint
can undergo elastic stretching together with the hairs. When a hair
detaches from the surface, the leg undergoes an elastic recoil due
to the stored elastic energy. Suppose the leg relaxes upward by a
recoil length, $\Delta l$. For $n$ hairs in contact, the force balance
before and after a hair detaches is given respectively by:

@@

We consider three possible mechanisms to detach the adhesive pad from
a surface: 1) \emph{Fixed pull}, where the pad is pulled vertically
up while keeping a fixed leg joint; 2) \emph{Free pull,} where the
pad is pulled vertically up while keeping the leg joint free to allow
rotation of the array; 3) \emph{Flex}, where the pad is hinged
to an external point (claw-hinge), and detached in a rotary fashion,
emulating the claw function in insects. To investigate each case in
detail, let us assume a pad with $N_{t}=25$ hairs and $\hat{f_{p}}=0.1$
attached to a stiff leg ($k_{l}\rightarrow\infty$). The situation
of a soft leg ($k_{l}/k_{h}=10$) is also considered for the first
two cases involving vertical detachment.

@@

\begin{figure}[H]
\includegraphics{Case1-Fixed_pull}\caption{\textbf{Detachment by fixed pull.} 
Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a fixed joint between the pad and the leg (\emph{Fixed pull}). The
tilt angle, $\theta$, of the array is kept fixed during detachment.
The leg is either stiff ($k_{l}\rightarrow\infty$) or soft ($k_{l}/k_{h}=10$).
Positive force values represent attraction. The green dashed line
represents the maximum possible adhesion for the pad. All values are
normalized to dimensionless forms, as described in text.\label{fig:Fixed-pull}}
\end{figure}

The fibrillar adhesive pad can be detached by pulling it vertically upwards
while maintaining a constant tilt angle, $\theta$, with the surface.
This can be achieved if the joint connecting the leg to the array is kept fixed. 
Equations \ref{eq:force_n_d}, \ref{eq:d_max}
and \ref{eq:simul_detach} can be used to get the resulting force-distance
curves.

@@

\includegraphics{Case2-Free_pull}\caption{\textbf{Detachment by free pull.} 
Force distance curves for a fibrillar adhesive pad, pulled vertically upwards with
a free joint between the pad and the leg (\emph{Free pull}). $s$
is the distance between the leg joint and the right end of the array
and $L$ is the array length. The free joint allows further tilting
of the array during the vertical pull. The leg is either stiff ($k_{l}\rightarrow\infty$)
or soft ($k_{l}/k_{h}=10$). Positive force values represent attraction.
The green dashed line represents the maximum possible adhesion for
the pad. All values are normalized to dimensionless forms, as described
in text. \label{fig:Free-pull}}

@@

Similar to the previous case, we once again consider the situation
where the adhesive pad is pulled vertically upwards for detachment.
However now, the leg joint is assumed to be free. In this case, the
array will reorient itself to maintain a zero net torque about the
joint during the entire detachment process. At any given instant,
the tilt angle, $\theta$, can be found by setting $M_{net}$ to zero
in equation \ref{eq:moment_n_d} to get:

@@

Using the above relation together with equations \ref{eq:force_n_d},
\ref{eq:d_max} and \ref{eq:simul_detach}, we can find force-distance
curves during a free vertical pull of the adhesive pad. Since the
position of the leg joint will influence the net torque, we use the
ratio, $s/L$, to study its effect on the detachment forces.

Maximum adhesion is seen when the leg joint is positioned at the centre
of the array, i.e. $s/L=0.5$ (Figure \ref{fig:Free-pull}). Here,
the net torque due to the hairs is balanced by symmetry and the array
remains parallel to the substrate until all hairs detach simultaneously
at $\hat{d_{s}}=0.1$. Shifting the position of the joint further
away from the array centre leads to lower forces or adhesion. The
resulting torque imbalance will tilt the array, which reduces the
net force due to the \emph{elastic weakening} effect, as described
in the previous section. Higher $s/L$ increases the net torque to
be balanced, leading to a higher tilt of the array and thus lower
net force.

@@

\begin{figure}[H]
\includegraphics{Case3-Fixed_flex}\caption{\textbf{Detachment by flex.}
Force curves for a fibrillar adhesive pad detached by flexing it about the claw.
$\hat{F}_{pull}=\frac{F_{pull}}{k_{h}w}$ is the normalised pulling
force necessary to apply the torque about the claw-hinge for detachment,
$\hat{F}_{hinge}=\frac{F_{hinge}}{k_{h}w}$ is the normalised reaction
force on the claw-hinge, $\hat{d_{s}}=\frac{d_{s}-l_{h,0}}{w}$ is
the normalized vertical distance of the claw-hinge from the surface.
Here, $s_{l}/L=1$ and $s_{h}/w=10$. The green dashed line represents
the maximum possible adhesion for the pad. \label{fig:Flex}}
\end{figure}

Instead of a vertical pull, the adhesive pad can also be detached
by rotating it about the claw-hinge, located outside the array. Such
a mode of detachment will be driven by a torque applied by the leg
to rotate the pad around the claw-hinge until all the hairs lose contact.
Let $s_{h}$ be the distance between the claw-hinge and the right
end of the array; $s_{l}$ be the distance between the fixed leg joint
and the right end of the array. To illustrate the mechanism, let us
fix $s_{l}/L=1$ and $s_{h}/w=10$ and vary the vertical claw-hinge
distance, $d_{s}$. At any particular instant, the pulling force applied
by the leg, $F_{pull}=M_{net}/\left(s_{l}+s_{h}\right)$, where $M_{net}$
is obtained by setting $s=-s_{h}$ in equation \ref{eq:moment_n_d}.
Equation \ref{eq:force_n_d} will give us the reaction force acting
on the claw-hinge, $F_{claw-hinge}$.

@@

The dimensionless parameter, $\hat{f}_{p}=\frac{f_{p}}{k_{h}w}$,
governs the strength and compliance of the array, where, high values
represent a dense array of strongly adhering soft hairs. Let us consider
the case of an adhesive pad with $N_{t}=25$ hairs and look at how
$\hat{f}_{p}$ influences the reduction factor under each mode of
detachment (Figure \ref{fig:Reduction-factor_fp}). From the colour plots, 
one can generally say that lower values of $\hat{f}_{p}$ increase
the reduction factor and thus provide a greater possibility of controlling
adhesion.

\begin{figure}[H]
\noindent \includegraphics{Reduction_factor-fp}\caption{\textbf{Effect of $\hat{f}_{p}$ on reduction factor.}
Colour plots showing the effect of the dimensionless parameter, $\hat{f}_{p}$,
on the reduction factor for each mode of detachment. Here, we fix
the number of hairs, $N_{t}=25$. All values are normalized to dimensionless
forms, as described in text. \label{fig:Reduction-factor_fp}}
\end{figure}

@@

For the case of detachment via \emph{free pull}, $\hat{f}_{p}$ has
no influence on the reduction factor. On the other hand, shifting
the position of the leg joint further away from the array (i.e. high
$s/L$) results in large values of $r$. In this scenario, the higher
torque exerted by the array leads to a higher tilt, and thus increases
the reduction factor via \emph{elastic weakening}, saturating to the
maximum value of 25.

@@

Let us now fix $\hat{f}_{p}=0.1$ and investigate the influence of
the number of hairs, $N_{t}$, on the reduction factor (Figure \ref{fig:Reduction-factor-Nt}).
The colour plots show that high $N_{t}$ increases $r$ irrespective
of the mode of detachment. Under a tilted state, more number of hairs
are compressed when $N_{t}$ is high, which reduces the net adhesion.
This highlights the advantage of having a split contact design found
in many biological systems. Higher number of hairs offer a better
control over adhesion and thus is more suited for reversible attachment
and detachment during locomotion. The specific trends of reduction
factor for each mode of detachment can be understood by similar arguments
of \emph{elastic weakening}, as discussed in the previous section.

\begin{figure}[H]
\noindent \begin{centering}
\includegraphics{Reduction_factor-Nt}\caption{\textbf{Effect of $N_{t}$ on reduction factor.} 
Colour plots showing the effect of the number of hairs, $N_{t}$, on
the reduction factor for each mode of detachment. Here, we fix the
dimensionless parameter, $\hat{f}_{p}=0.1$. All values are normalized
to dimensionless forms, as described in text. \label{fig:Reduction-factor-Nt}}
\par\end{centering}
\end{figure}

@@

Based on the above assumptions, we can come up with force-distance
curves to detach the beetle's leg via various pathways (Figure \ref{fig:Detachment-pathways}).
First, let us assume the joint H\textsubscript{2} to be fixed, such
that both the distal and proximal pads can be combined to behave like a single
long pad with $N_{t}=60$ hairs. Path 1 shows the case where the pad
shows maximum possible adhesion. Here, the combined pad is vertically
pulled upwards while keeping the array perfectly parallel to the surface.
If this combined pad is detached by keeping H\textsubscript{1} fixed and maintaining
a tilt of $1\lyxmathsym{\textdegree}$ with the surface (path 2),
the forces dramatically reduces, with around 10 times reduction in
the adhesion compared to path 1. We can also detach the pad by switching
between the different mechanisms. Path 3 shows one such example, where,
initially the leg is pulled vertically up while keeping H\textsubscript{1}
fixed, stretching the hairs. On reaching point $a$, H\textsubscript{1}
is set free, which results in a sudden drop in force due to the excess
torque by the stretched hairs, tilting the array. Beyond this, the
force curve follows the \emph{free pull} mechanism, with $\thicksim3.5$
times reduction in adhesion. An alternate strategy of switching between
mechanisms would be to first apply a load on the pad (path 4) and
compress the hairs until point $b$. Beyond this point, the claw can
be used as a hinge to detach the pad via flexing it around H\textsubscript{3},
which once again reduces the adhesion force. Now, if we assume the
joint H\textsubscript{2} to be free such that the two pads can behave
distinctly, we can consider the scenario where the proximal pad is
flexed around the distal pad at H\textsubscript{2} while keeping
H\textsubscript{1} fixed (path 5). After the proximal pad has completely
detached, H\textsubscript{1} can be freed up at point $c$ to detach
the distal pad via \emph{free pull} with very little force. This pathway
results in a $\thicksim5$ times reduction in adhesion.

\begin{figure}[H]
\includegraphics{Detachment_pathways}\caption{\textbf{Beetle leg detachment pathways.}
Force distance curves showing possible detachment pathways for a dock
beetle's leg, derived from the spring contact model. The curves are
offset laterally for clarity. Colours represent the distinct detachment
pathways, as labelled correspondingly, with arrows indicating the
direction of retraction. The different line style denotes the specific
detachment mechanism followed by any region of the pathway. The inset
schematic shows the assumed locations of the different hinges (H\protect\textsubscript{1},
H\protect\textsubscript{2} and H\protect\textsubscript{3}) employed
by the leg.\label{fig:Detachment-pathways}}
\end{figure}

The above analysis illustrates how the design of the beetle's hairy
adhesive pads is suitable for modulating its adhesion. Effective control
and release of its leg joints can help the insect to reduce the pad's
adhesion, allowing it to detach with little effort. High reduction
in adhesion is seen when the pad is tilted relative to the surface
during detachment, as a result of \emph{elastic weakening}. To the
best of our knowledge, there is no direct experimental evidence that
beetles can modulate its adhesion by taking advantage of this effect.
Considering that hair deformation occurs at length scales below 10
\textgreek{m}m, direct observation of this effect on running beetles
would be challenging. A recent study on PDMS micro-pillar arrays,
however, does indeed show a strong reduction in the adhesion force
due to slight misalignments with the surface \cite{RN252}. Based on
previously reported microscopic investigation of freely walking dock
beetles \cite{RN83}, we argue that the following experimental observations
provide support to our proposed theory: 1) The detachment was shown
to follow a three-dimensional twist of the leg, which suggests a complex
inhomogeneous deformation of hairs across the array, leading to \emph{elastic
weakening}, which is suited for easy detachment. 2) The beetle can at times instantaneously detach all
its legs and ``drop'' itself while upside down. This could be explained
by the beetle freeing up its leg joints and using just its body weight
to provide the necessary force to detach all its legs via \emph{free
pull} (similar to path 3 above). 3) Only a fraction of the beetle's
pads made contact with the surface during locomotion, which indicates
that the pads should naturally be in a slightly tilted state. This
not only reduces the contact area, but also non-uniformly deforms
the hairs, both leading to a reduction in adhesion for easy detachment.
4) Contact images showed that the array ``peels'' from the proximal
to distal direction during detachment. However, the beetle's hairs
are attached to a relatively stiff backing \cite{RN53}, so it wouldn't
be able to ``peel'' its array, since peeling, strictly speaking,
depends on the elastic contribution of a thin flexible backing as
it bends during the process \cite{RN147}. Rather, the ``peeling''
observed in the beetle's case should be a result of the pad detaching
from the surface in a tilted orientation, causing the hairs to distinctly detach
in sequence. 5) The time scale of detachment was reported to be an
order of magnitude shorter than the attachment time scale, which could
be a result of the elastic recoil of the leg causing a spontaneous
propagation of the detachment front (Figure \ref{fig:Fixed-pull}).

There exists a limit to how much the pad can tilt, depending on its
geometry and material properties. Suppose the hair has a maximum linear
elastic strain limit, $\varepsilon_{m}$, and natural length, $l_{h,0}$.
Based on Figure \ref{fig:Schematic}, if the right most hair is compressed
to its elastic limit, one can derive from simple geometry, that, the
corresponding maximum limit in tilt angle is given by: 
\[
\theta_{limit}=\arctan\frac{l_{h,0}\varepsilon_{m}}{\left(N_{t}-1\right)w}
\]
$\theta_{limit}$ will limit the reduction factor for each of the
detachment mechanisms presented. Longer hairs can result in a lateral
collapse or bundling of hairs, imposing an additional constraint on
$\theta_{limit}$. Large deformation of hairs can also lead to buckling,
which will further limit the reduction in adhesion due to the smaller
effective modulus. Thus, the geometry of the individual hairs is also
crucial in effectively reducing adhesion by \emph{elastic weakening}.
Ideally, hairs with large diameters would be preferable to eliminate
buckling, and optimally long hairs would provide a large range of
angles to which the pad can be tilted while avoiding bundling.

@@

The net torque about the joint due to the deformed hairs of the array
is,

@@

Here, $\lambda_{i}=s-\left(\frac{n-1}{2}w-\frac{x_{i}}{\cos\theta}\right)$
is the length of the lever arm between the $i^{th}$ hair and the
joint.
